---
title: "Binomial!"
output: 
  html_document:
    theme: cerulean
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
require(ggplot2)
require(dplyr)
require(tidyr)
require(learnr)
source("https://raw.githubusercontent.com/kbodwin/Introductory-Statistics-Labs/master/Scripts/plotDensities.R")
source("https://raw.githubusercontent.com/kbodwin/Introductory-Statistics-Labs/master/Scripts/makeStrings.R")
titanic = read.csv('https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Datasets/titanic.csv')
titanic <- titanic %>% mutate(
  Survived = factor(Survived,
                    levels = c(0,1),
                    labels = c("Died", "Lived")),
  Passenger.Class = factor(Pclass, 
                           levels = c(1,2,3), 
                           labels = c("First", "Second", "Third"))
)
correct <- "<font color='green'>Correct!</font><br><br>"
incorrect <- "<font color='red'>Sorry, try again.</font><br><br>"
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
checker <- function(
  label,
  user_code,
  solution_code,
  check_code,
  envir_result,
  evaluate_result,
  envir_prep,
  ...
) {
  list(message = "Great job!", correct = TRUE, location = "append")
}
``` 

# {.tabset}

## Intro {.tabset .tabset-pills}

### Purpose 
This applet will go over three funtions to help you with the binomial process: <br>
**dbinom** will calculate the probability of a given number of successes <br>
**pbinom** will calculate the probability of less than or equal to a given number of successes <br>
**qbinom** will return the value of the inverse cumulative density function, this means you can use it to find the $p^{th}$ quantile of the binomial distribution <br>

### What is a binomial experiment?
A binomial experiment is an experiment satisfying the following conditions. 

1.	There is a fixed number of trials, denoted by n.
2.	A trial has only two possible outcomes: a “success” or a “failure.”
3.	The trials of the experiment are independent of each other (if the outcome of one trial is a “success,” then the chance that another outcome is a “success” is not influenced).
4.	The probability of observing a “success” on each trial remains the same, denoted by p. We denote the probability of a “failure” by q = 1 – p. 

In general, a binomial random variable is a random variable that represents the number of “successes” in n trials.

If a random variable follows a binomial distribution we write it as: `X~Binomial(n, p)`




## dbinom {.tabset .tabset-pills}

### Create your own binomial distribution!

The first important function is `dbinom()`, which finds the *probability* of some number of successes.

<div class = "row">

<div class = "col-md-4">
```{r factor_var, out.width=c('50%', '50%'), fig.show='hold'}
textInput("n",
          "How many trials do you have?",
          value = "0",
          width = '80%', placeholder = NULL)
textInput("p",
          "What is the probability of success at each trial?",
          value = "0",
          width = '80%', placeholder = NULL)
textInput("k",
          "We are interested in the probability of how many successes?",
          value = "0",
          width = '80%', placeholder = NULL)
bs_1_1 = reactive("dbinom(%ni, size = %ni, prob = %ni)")

```

</div>

<div class = "col-md-6">
```{r}
plotOutput("distn_1")
# the number of successes doesn't show up as part of the plot
# slightly weird space to the left of plot
```
</div>

<div class = "col">
Random variable and probability statement:
```{r}
htmlOutput("binom_var_1")
```

Code:
```{r}
htmlOutput("code_1_1")
```

Result:
```{r}
verbatimTextOutput("output_1_1")
```
</div>
</div>

```{r, context="server"}

output$code_1_1 <- renderText(
        makePrintText(
          base_string = bs_1_1(),
          old_input = c(),
          new_input = c(input$k, input$n, input$p)
        )
  )
output$output_1_1 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_1_1(), 
      old_input = c(),
      new_input = c(input$k, input$n, input$p)
    )
   ))
})
output$binom_var_1 <- renderText(
        makePrintText(
          base_string = "X ~ Binomial(%ni, %ni) \n P(X = %ni) = ?",
          old_input = c(),
          new_input = c(input$n, input$p, input$k)
        )
  )
output$distn_1 <- renderPlot(
  illustrate_binom(input$n, input$p, x = input$k, equal_to = TRUE)
)
```


### Practice Problems

In the boxes below each question, you may run any line of **R** code. Use the consoles to answer the following questions and click **Submit Answer** to check your code.

I flip 30 fair coins.  What is the probability that I get exactly 15 tails?
```{r 1a, exercise = TRUE, exercise.checker = TRUE, exercise.lines = 3}

```

```{r 1a-check}
dbinom(15, 30, .5)
checker()
```

16% of all Starburst candies in a pack are cherry flavored.  What is the probability that in a pack of 12, I get exactly 4 cherry Starbursts?
```{r 1b, exercise = TRUE, exercise.checker = TRUE, exercise.lines = 3}

```

```{r 1b-check}
dbinom(4, 12, 0.16)
checker()
```

In an upcoming election, 40% of voters prefer Candidate A, and 60% prefer Candidate B.  If only 16 people vote, what is the probability that the vote is NOT tied?
```{r 1c, exercise = TRUE, exercise.checker = TRUE, exercise.lines = 3}

```

```{r 1c-check}
1- dbinom(8, 16, 0.4)
checker()
```

Let's take a look at the `titanic` data set. In this data set, each observation is a passenger who was on the Titanic; there are a number of variables including passenger class, age, sex, whether the passenger survived or not. If I randomly selected 60 passengers, what is the probability that exactly one third of them survived? (Hint: This question takes a little bit of work, think about how you would go about finding the probability of people who survived. Still confused? See the Quantitative Variables lab exercise.)
```{r 1d, exercise = TRUE, exercise.checker = TRUE, exercise.lines = 3}

```

```{r 1d-check}
dbinom(20, 60, 342/887)
checker()
```


## pbinom {.tabset .tabset-pills}

### What does pbinom look like?

It is not very common that we are interested in the probability of some exact number of successes.
More often, we want to quantify **evidence** by finding the probability of some number **or more** of successes.  For this, we will use *pbinom()*.  Note that by default, `pbinom()` finds the probability of getting **less than or equal to** the given number. Try it below: <br><br>
<div class = "row">
<div class = "col-md-4">
```{r pbinom2}
textInput("n2",
          "How many trials do you have?",
          value = "0",
          width = '80%', placeholder = NULL)
textInput("p2",
          "What is the probability of success at each trial?",
          value = "0",
          width = '80%', placeholder = NULL)
textInput("k2",
          "We are interested in the probability of less than or equal to how many successes?",
          value = "0",
          width = '80%', placeholder = NULL)
bs_1_2 = reactive("pbinom(%ni, size = %ni, prob = %ni)")
```
</div>
<div class = "col-md-6">
```{r}
plotOutput("distn_2")
```
</div>
<div class = "col">
Random variable and probability statement:
```{r}
htmlOutput("binom_var_2")
```

Code:
```{r}
htmlOutput("code_2")
```

Result:
```{r}
verbatimTextOutput("output_2")
```
</div>
</div>



```{r, context="server"}
output$code_2 <- renderText(
        makePrintText(
          base_string = bs_1_2(),
          old_input = c(),
          new_input = c(input$k2, input$n2, input$p2)
        )
  )
output$output_2 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_1_2(), 
      old_input = c(),
      new_input = c(input$k2, input$n2, input$p2)
    )
   ))
})
output$binom_var_2 <- renderText(
        makePrintText(
          base_string = "X ~ Binomial(%ni, %ni) \n P(X ≤ %ni) = ?",
          old_input = c(),
          new_input = c(input$n2, input$p2, input$k2)
        )
  )
output$distn_2 <- renderPlot(
  illustrate_binom(input$n2, input$p2, x = input$k2)
)
```


### Practice Problems
Use the consoles below to answer each question and click **Submit Answer** to check your code. 

I flip 30 fair coins.  What is the probability that I get less than half tails?
```{r 2a, exercise = TRUE, exercise.checker = TRUE, exercise.lines = 3}

```

```{r 2a-check}
pbinom(14, 30, .5)
checker()
```

16% of all Starburst candies in a pack are cherry flavored.  What is the probability that in a pack of 12, I get more than 4 cherry Starbursts?
```{r 2b, exercise = TRUE, exercise.checker = TRUE, exercise.lines = 3}

```

```{r 2b-check}
1-pbinom(4, 12, 0.16)
checker()
```

About 10% of all humans are left-handed.  If I randomly choose 43 humans, what is the probability that at least 8 of them are left handed?
```{r 2c, exercise = TRUE, exercise.checker = TRUE, exercise.lines = 3}

```

```{r 2c-check}
1- pbinom(7, 43, 0.1)
checker()
```

Back to our titanic data set! If you take a sample of 21 passengers, what is the probability that at least 13 of them were in first class?
```{r 2d, exercise = TRUE, exercise.checker = TRUE, exercise.lines = 3}

```

```{r 2d-check}
1- pbinom(12, 21, 216/887)
checker()
```

By default, `pbinom()` tells you the probability of getting *less than or equal to* the given number of successes. For problems like the last one in this set of questions, that means you had to subtract from 1 to get the probability you wanted (which was the probability of the given number of successes *or more*). However, you could also add `lower.tail = FALSE` into the `pbinom()` function and it will do the same thing. (Note that without using `lower.tail = FALSE`, you are asking for the probability of less than or equal to a given number and when you do include `lower.tail = FALSE`, you are asking for the probability of greater than a given number.)

```{r lower_tail, exercise = TRUE, exercise.checker = TRUE, exercise.lines = 3}

```

```{r lower_tail-check}
pbinom(12, 21, 216/887, lower.tail = F)
checker()
```


## qbinom {.tabset .tabset-pills}

### What does qbinom look like?
Finally, sometimes we want to ask probability questions "in reverse".  For example, we might ask "If I flip 100 coins, I can be 95% sure that I will get at least how many tails?".  We could answer this via guess-and-check with `pbinom()`, but fortunately, there is a function for it in **R** called `qbinom()`. <br><br>

<div class = "row">
<div class = "col-md-4">
```{r qbinom}
textInput("n3",
          "How many trials do you have?",
          value = "0",
          width = '80%', placeholder = NULL)
textInput("p3",
          "What is the probability of success at each trial?",
          value = "0",
          width = '80%', placeholder = NULL)
textInput("q",
          "We want to find the number of successes associated with the _____th quantile (in decimal form).",
          value = "0",
          width = '80%', placeholder = NULL)
bs_3 = reactive("qbinom(%ni, size = %ni, prob = %ni)")
```
</div>
<div class = "col-md-6">
```{r}
plotOutput("distn_3")
```
</div>
<div class = "col">
Random variable and probability statement:
```{r}
htmlOutput("binom_var_3")
```

Code:
```{r}
htmlOutput("code_3")
```

Result:
```{r}
verbatimTextOutput("output_3")
```
</div>
</div>


```{r, context="server"}
output$code_3 <- renderText(
        makePrintText(
          base_string = bs_3(),
          old_input = c(),
          new_input = c(input$q, input$n3, input$p3)
        )
  )
output$output_3 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_3(), 
      old_input = c(),
      new_input = c(input$q, input$n3, input$p3)
    )
   ))
})
output$binom_var_3 <- renderText(
        makePrintText(
          base_string = "X ~ Binomial(%ni, %ni) \n P(X ≤ ?) = %ni",
          old_input = c(),
          new_input = c(input$n3, input$p3, input$q)
        )
  )
output$distn_3 <- renderPlot(
  illustrate_binom(input$n3, input$p3, q = input$q)
)
```


### Practice Problems
Use the consoles below to answer each question and click **Submit Answer** to check your code. 

I flip 30 fair coins.  There is a 90% chance that I get less than how many heads?
```{r 3a, exercise = TRUE, exercise.checker = TRUE, exercise.lines = 3}

```

```{r 3a-check}
qbinom(0.9, 30, .5)
checker()
```

16% of all Starburst candies in a pack are cherry flavored.  I can be 85% sure that I will get at least how many cherry flavored Starbursts in a pack of 12?
```{r 3b, exercise = TRUE, exercise.checker = TRUE, exercise.lines = 3}

```

```{r 3b-check}
qbinom(.15, 12, 0.16)
checker()
```

About 10% of all humans are left-handed.  I think there is a 95% chance that fewer than how many of my 35 students are left-handed?
```{r 3c, exercise = TRUE, exercise.checker = TRUE, exercise.lines = 3}

```

```{r 3c-check}
qbinom(.95, 35, 0.1)
checker()
```


## Extra Problems {.tabset .tabset-pills}

A good habit to get into when coding is to "translate" your code to English in your head, and try to predict what the output will be. Are you truly asking the computer to calculate what you hoped? (Remember: computers will almost always output *something*! It is your responsibility as the scientist to make sure you have asked the right question.)

For the following, try to guess what the output would be.  Do **not** run the code; instead, try to read it as a probability question and think through the answer.

*Give answers to 2 decimals*

```{r q_4}
textInput("a_4.1",
           "dbinom(1, size = 1, prob = 0.7)",
           value = "",
           width = '80%', placeholder = NULL)
htmlOutput("check_4.1")
textInput("a_4.2",
           "pbinom(60, size = 60, prob = 0.13)",
           value = "",
           width = '80%', placeholder = NULL)
htmlOutput("check_4.2")
textInput("a_4.4",
           "qbinom(1, size = 60, prob = 0.13)",
           value = "",
           width = '80%', placeholder = NULL)
htmlOutput("check_4.4")
textInput("a_4.3",
           "pbinom(50, size = 100, prob = 0.5)",
           value = "",
           width = '80%', placeholder = NULL)
htmlOutput("check_4.3")
textInput("a_4.5",
           "qbinom(0.5, size = 100, prob = 0.5)",
           value = "",
           width = '80%', placeholder = NULL)
htmlOutput("check_4.5")
```

```{r}
checkthis = function(input, solution){
  if(input == ""){
     ""
  }else if(as.numeric(input) == solution){
      correct
  }else{
    incorrect
  }
}
```

```{r, context = "server"}
output$check_4.1 <- renderText(checkthis(input$a_4.1, 0.7))
output$check_4.2 <- renderText(checkthis(input$a_4.2, 1))
output$check_4.3 <- renderText(checkthis(input$a_4.3, 0.5))
output$check_4.4 <- renderText(checkthis(input$a_4.4, 60))
output$check_4.5 <- renderText(checkthis(input$a_4.5, 150))
```
