---
title: "Practice Binom"
output: 
  html_document:
    theme: cerulean
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
require(ggplot2)
require(dplyr)
require(tidyr)
require(learnr)
source("https://raw.githubusercontent.com/kbodwin/Introductory-Statistics-Labs/master/Scripts/plotDensities.R")
source("https://raw.githubusercontent.com/kbodwin/Introductory-Statistics-Labs/master/Scripts/makeStrings.R")
titanic = read.csv('https://raw.githubusercontent.com/kbodwin/ShinyLabs/master/Datasets/titanic.csv')
titanic <- titanic %>% mutate(
  Survived = factor(Survived,
                    levels = c(0,1),
                    labels = c("Died", "Lived")),
  Passenger.Class = factor(Pclass, 
                           levels = c(1,2,3), 
                           labels = c("First", "Second", "Third"))
)
knitr::opts_chunk$set(echo = FALSE)
```
 

# {.tabset}

## Intro {.tabset .tabset-pills}

A binomial experiment is an experiment satisfying the following conditions. 

1.	There is a fixed number of trials, denoted by n.
2.	A trial has only two possible outcomes: a “success” or a “failure.”
3.	The trials of the experiment are independent of each other (if the outcome of one trial is a “success,” then the chance that another outcome is a “success” is not influenced).
4.	The probability of observing a “success” on each trial remains the same, denoted by p. We denote the probability of a “failure” by q = 1 – p. 

In general, a binomial random variable is a random variable that represents the number of “successes” in n trials.

If a random variable follows a binomial distribution we write it as: X~Binomial(n, p)

This applet will go over three funtions to help you with the binomial process: <br>
**dbinom** will calculate the probability of a given number of successes <br>
**pbinom** will calculate the probability of less than or equal to a given number of successes <br>
**qbinom** lets us work our way backward to find the number of successes for a given probability <br>
^ should fix qbinom to mention that it's that it's a range of successes
*add binomial example*

## dbinom {.tabset .tabset-pills}

### Create your own binomial distribution!

The first important function is `dbinom()`, which finds the *probability* of some number of successes.

```{r factor_var}
textInput("n",
          "How many trials do you have?",
          value = "0",
          width = '80%', placeholder = NULL)
textInput("p",
          "What is the probability of success at each trial?",
          value = "0",
          width = '80%', placeholder = NULL)
textInput("k",
          "We are interested in the probability of how many successes?",
          value = "0",
          width = '80%', placeholder = NULL)
bs_1_1 = reactive("dbinom(%ni, size = %ni, prob = %ni)")
```

```{r, context="server"}
# nothing happens here, you can put in the input but rmd doesn't do anything with it
#output$plot = renderPlot(barplot(dbinom(0:input$n, size = input$n, prob = input$p)))
output$code_1_1 <- renderText(
        makePrintText(
          base_string = bs_1_1(),
          old_input = c(),
          new_input = c(input$k, input$n, input$p)
        )
  )
output$output_1_1 <- renderPrint({
 eval(parse(text = 
    makeEvalText(
      base_string = bs_1_1(), 
      old_input = c(),
      new_input = c(input$k, input$n, input$p)
    )
   ))
})
output$binom_var_1 <- renderText(
        makePrintText(
          base_string = "X ~ Binomial(%ni, %ni) \n P(X = %ni) = ?",
          old_input = c(),
          new_input = c(input$n, input$p, input$k)
        )
  )
output$distn_1 <- renderPlot(
  illustrate_binom(input$n, input$p, x = input$k, equal_to = TRUE)
)
```


### practice problems

In the box below, you may run any line of **R** code.  Use the box, and/or the exercise above, to answer the following questions.

```{r dbinom, exercise = TRUE, exercise.lines = 5}
```

```{r q_1}
textInput("a_1.1",
           "I flip 30 fair coins.  What is the probability that I get exactly 15 tails?",
           value = "",
           width = '80%', placeholder = NULL)
htmlOutput("check_1.1")
textInput("a_1.2",
           "16% of all Starburst candies in a pack are cherry flavored.  What is the probability that in a pack of 12, I get exactly 4 cherry Starbursts?",
           value = "",
           width = '80%', placeholder = NULL)
htmlOutput("check_1.2")
textInput("a_1.3",
           "In an upcoming election, 40% of voters prefer Candidate A, and 60% prefer Candidate B.  If only 16 people vote, what is the probability that the vote is NOT tied?",
           value = "",
           width = '80%', placeholder = NULL)
htmlOutput("check_1.3")
textInput("a_1.4",
           "Let's take a look at the `titanic` data set. In this data set, each observation is a passenger who was on the Titanic; there are a number of variables including passenger class, age, sex, whether the passenger survived or not. If I randomly selected 60 passengers, what is the probability that exactly one third of them survived? (Hint: This question takes a little bit of work, think about how you would go about finding the probability of people who survived. Still confused? See the Quantitative Variables lab exercise.)",
           value = "",
           width = '80%', placeholder = NULL)
htmlOutput("check_1.4")
```

```{r, context = "server"}
output$check_1.1 <- renderText(
  if(input$a_1.1 == ""){
     ""
  }else if(round(input$a_1.1,2) == round(dbinom(15, 30, .5), 2)){
      correct
  }else{
    incorrect
  }
)
output$check_1.2 <- renderText(
  if(input$a_1.2 == ""){
     ""
  }else if(round(input$a_1.2,2) == round(dbinom(4, 12, 0.16), 2)){
      correct
  }else{
    incorrect
  }
)
output$check_1.3 <- renderText(
  if(input$a_1.3 == ""){
     ""
  }else if(round(input$a_1.3,2) == round(1- dbinom(8, 16, 0.4), 2)){
      correct
  }else{
    incorrect
  }
)
output$check_1.4 <- renderText(
  if(input$a_1.4 == ""){
     ""
  }else if(round(input$a_1.1,2) == round(dbinom(20, 60, 342/887), 2)){
      correct
  }else{
    incorrect
  }
)
```

## pbinom {.tabset .tabset-pills}

### practice creating
The first important function is `dbinom()`, which finds the *probability* of some number of successes.
some input stuff

Random variable and probability statement:

Code:

Result:

Illustration:


### practice problems

some stuff


## qbinom {.tabset .tabset-pills}

### practice creating
The first important function is `dbinom()`, which finds the *probability* of some number of successes.
some input stuff

Random variable and probability statement:

Code:

Result:

Illustration:


### practice problems

some stuff
